---
title: "Plots for Smith Majors Project"
author: "Camila Maldonado Franco"
output:
  pdf_document:
    latex_engine: xelatex
  html_document: default
---

-Starting to use ggplot to create some visualizations, yet first I will need to group by and count the majors in each year for each department. 
Some ideas for visualizations are: 
  -A linegraph to see the relationship between number of majors throughout the years in each department.
  -A barplot to check the total number of majors in each department
  


```{r}
table4 <- table4 %>% 
  rename(Department = Dept.) 

students_per_dept <- table4 %>%
  pivot_longer(cols = `2008-09`:`2020-21`, names_to = "year", values_to = "num_majors") %>%
  group_by(Department, year) %>%
  summarize(total_students = sum(num_majors)) 
```


```{r}

majors_dept <- students_per_dept %>%
  group_by(Department) %>% 
  summarize(total = sum(total_students))

ggplot(majors_dept, aes(total, reorder(Department, total), fill = Department)) + geom_col() + scale_fill_manual(values = c("#CD5C5C",
"#F08080",
"#E9967A",
"#FA8072",
"#FFA07A",
"#FF4500",
"#FF8C00",
"#FFA500",
"#FFD700",
"#B8860B",
"#DAA520",
"#EEE8AA",
"#BDB76B",
"#F0E68C",
"#808000",
"#FFFF00",
"#9ACD32",
"#556B2F",
"#6B8E23",
"#7CFC00",
"#7FFF00",
"#ADFF2F",
"#006400",
"#008000",
"#228B22",
"#00FF00",
"#32CD32",
"#90EE90",
"#98FB98",
"#8FBC8F",
"#00FA9A",
"#00FF7F",
"#2E8B57",
"#66CDAA",
"#3CB371",
"#20B2AA",
"#2F4F4F",
"#008080",
"#008B8B",
"#00FFFF",
"#00FFFF",
"#E0FFFF",
"#00CED1",
"#40E0D0",
"#48D1CC")) + labs(title = "Smithies favorite majors", x = "Count", y = "Departments") 

```


```{r}
sds <- students_per_dept %>% 
  filter(Department == "SDS") %>% 
  group_by(Department)

ggplot(sds, aes(year, total_students)) + geom_line(group = 1) + labs(title = "Statistical and Data Science Majors", x = "Year", y = "Total number of Students" )

```


```{r}
edc <- students_per_dept %>% 
  filter(Department == "EDC") %>% 
  group_by(Department)

ggplot(edc, aes(year, total_students)) + geom_line(group = 1) + labs(title = "Education & Child Studies Majors", x = "Year", y = "Total number of Students" )
```


-Spaguetti graph 

```{r}
linegraph2 <- students_per_dept %>% 
  mutate(year = as.Date(year, format = "%Y"))

ggplot(linegraph2, aes(year, total_students, color = Department)) + geom_line()  + scale_color_viridis(discrete = TRUE, option = "C") + scale_x_date(date_breaks = "1 year", date_labels = "%Y") + labs(title = "Graduates and their majors throughout time", x = "Year", y = "Total number of Students" )



```


-Loading library Ploty which is used to make interactive visualizations in R. 

```{r}
library(plotly)
```


```{r}
#Trying an animated linegraph 
plot <- linegraph2 %>%
ggplotly(x = ~year, y = ~total_students,
split = ~Department,
frame = ~frame,
type = 'scatter',
mode = 'lines',
line = list(simplify = F))

plot <- linegraph2 %>% layout(
xaxis = list(
title = "Date",
zeroline = F
),
yaxis = list(
title = "Total Number of Students",
zeroline = F))

plot <- linegraph2 %>% animation_opts(
frame = 100,
transition = 0,
redraw = FALSE)

#plot <- linegraph %>% animation_slider(
#hide = T)

#plot <- linegraph %>% animation_button(
#x = 1, xanchor = "right", y = 0, yanchor = "bottom")

```


-Filling out blanks in Division column 

```{r}
table5 <- for (i in 1:nrow(table4)) {
  if (table4$Division.[i] == "Division II - Natural Science") {
    table4$Division.[i] <- table4$Division.[i-1]
  }
}
```


